---
title: "Data Science Final Project"
subtitle: "MODEL 1B: SVM"
author: "Kjay O. Coca"
date: "2022-12-14"
output: pdf_document
---


```{r}
# Helper packages
library(dplyr)    # for data wrangling
library(ggplot2)  # for awesome graphics
library(rsample)  # for data splitting
library(readr)

# Modeling packages
library(caret)    # for classification and regression training
library(kernlab)  # for fitting SVMs
library(modeldata) #for Failure.binary data
library(forcats)

# Model interpretability packages
library(pdp)      # for partial dependence plots, etc.
library(vip)      # for variable importance plots
```

```{r}
set.seed(123)
radiomics_data <- read_csv("D:/1 MASTERS/STAT225/FINAL PROJECT/STAT 325 _FINAL PROJECT_/Normalize Radiomics Data.csv")
```

## Load Failure.binary data
```{r}
radiomics_data$Failure.binary=as.factor(radiomics_data$Failure.binary)
```

## SPLITTING FOR TRAINING AND TESTING
```{r}
set.seed(123)  # for reproducibility
for_split <- initial_split(radiomics_data, prop = 0.8, strata = "Failure.binary")
radiomicsdata_train <- training(for_split)
radiomicsdata_test  <- testing(for_split)
```
In this case, I set 80 percent for training data and 20 percent for testing data. There are 39 observation for testing and 158 observation for training and both have 413 variables.


## Linear
```{r}
caret::getModelInfo("svmLinear")$svmLinear$parameters
```

## Polynomial kernel
```{r}
caret::getModelInfo("svmPoly")$svmPoly$parameters
```

## Radial basis kernel
```{r}
caret::getModelInfo("svmRadial")$svmRadial$parameters
```

### Run SVM Model in Training phase
```{r, warning=FALSE}
set.seed(1854) 
svm_split <- train(
  Failure.binary ~ ., 
  data = radiomicsdata_train,
  method = "svmRadial",               
  preProcess = c("center", "scale"),  
  trControl = trainControl(method = "cv", number = 10),
  tuneLength = 10
)
```
Using `split_train()` function, tuning SVM model with radial basis kernel. Using `split_train()` function, we can tune an SVM model with radial basis kernel.

## Plot results
```{r}
ggplot(svm_split) + theme_light()
```


## Print results
```{r}
svm_split$results
```


## Control parameters for SVM
```{r}
class.weights = c("No" = 1, "Yes" = 10)
ctrl <- trainControl(
  method = "cv", 
  number = 10, 
  classProbs = TRUE,                 
  summaryFunction = twoClassSummary  
)

radiomicsdata_train$Failure.binary <- fct_recode(radiomicsdata_train$Failure.binary,No="0",Yes="1")

```


## Print the AUC values during Training
```{r}
# Tune an SVM
set.seed(123)  # for reproducibility
train_svm_auc <- train(
  Failure.binary ~ ., 
  data = radiomicsdata_train,
  method = "svmRadial",               
  preProcess = c("center", "scale"),  
  metric = "ROC",  # area under ROC curve (AUC)       
  trControl = ctrl,
  tuneLength = 10
)

# Print results
train_svm_auc$results
confusionMatrix(train_svm_auc)
```
The average accuracy of the trained model is 0.828 or 82.8 percent.


## Print the Top 20 important features during Training
```{r}
prob_yes <- function(object, newdata) {
  predict(object, newdata = newdata, type = "prob")[, "Yes"]
}

# Variable importance plot
set.seed(2827)  # for reproducibility
vip(train_svm_auc, method = "permute", nsim = 5, train = radiomicsdata_train, 
    target = "Failure.binary", metric = "auc", reference_class = "Yes", 
    pred_wrapper = prob_yes)
```

This are the top 20 important variable during Training. Failure variable is the most important. And next is Entrophy_cooc.W.ADC.


```{r}
features <- setdiff(names(radiomics_data), names(radiomics_data)[c(1,2)])
pdps <- lapply(features, function(x) {
  partial(train_svm_auc, pred.var = x, which.class = 2,  
          prob = TRUE, plot = TRUE, plot.engine = "ggplot2") +
    coord_flip()
})

grid.arrange(grobs = pdps,  ncol = 2)
```

## Print the AUC values during Testing

```{r}
radiomicsdata_test$Failure.binary=fct_recode(radiomicsdata_test$Failure.binary,No="0",Yes="1")

# Tune an SVM with radial 
set.seed(5628)  # for reproducibility
test_svm_auc <- train(
  Failure.binary ~ ., 
  data = radiomicsdata_test,
  method = "svmRadial",               
  preProcess = c("center", "scale"),  
  metric = "ROC",  # area under ROC curve (AUC)       
  trControl = ctrl,
  tuneLength = 10
)

# Print results
test_svm_auc$results
confusionMatrix(test_svm_auc)
```
The accuracy of test data in this model is 0.625 or just 62.5 percent. It smaller compare to trained data.


